---
title: Seurat Analysis of HNSCC data set
---
##LIBRARIES
Load the libraries:

```{r libraries}
library(limma)
library(dplyr)
library(Seurat)
library(patchwork)
```
##DATASET
Load the data set:
```{r data}
hnscc_data <- read.table(file='/home/ressf/Documenti/RessBachelorsThesisCode/Downstream_analysis/HNSCC/HNSCC_all_data.txt' , row.names=1, skip=6, sep='\t')
```
Create a Seurat Object:
```{r Seuratobj}
hnscc <- CreateSeuratObject(counts=hnscc_data)
```
Print the object to make sure everything is ok:
```{r printobj}
print(hnscc)
```
Violin plot of the data:
```{r plot1}
VlnPlot(hnscc, features=c("nFeature_RNA","nCount_RNA"),ncol=2)
```
Calculate percentage.mt and plot it:
```{r percentage}
hnscc[["percent.mt"]] <- PercentageFeatureSet(hnscc, pattern = "^MT-")
VlnPlot(hnscc, features=c("nFeature_RNA","nCount_RNA","percent.mt"),ncol=3)
```
##Normalization
Normalize the Data:
```{r normalization}
#normalization
hnscc_norm <- NormalizeData(hnscc, normalization.method="LogNormalize", scale.factor=10000)
```
#Feature selection
The data was normalized, we proceed:

```{r featureselection}
#feature selection
hnscc_norm <- FindVariableFeatures(hnscc_norm, selection.method="vst", nfeatures= 5902)
```
Select top 10 genes and label the 10 top expressed genes, plot it
```{r top10}
top10 <- head(VariableFeatures(hnscc_norm), 10)
plot1 <-VariableFeaturePlot(hnscc_norm)
plot2 <-LabelPoints(plot= plot1, points=top10, repel =TRUE)
plot1+plot2
```
Find Variable Features of the data:
```{r variablefeatures}
hnsccFeatures <- FindVariableFeatures(hnscc, selection.method="vst", nfeatures=5902)
```
#Scaling and PCA
Scaling the data:
```{r scaling}
# scaling the data
all.genes <-rownames(hnsccFeatures)
hnsccScaled<-ScaleData(hnscc_norm,features=all.genes)
```
Perform PCA:
```{r pca}
#perform linear dimensional reduction
hnsccScaled<- RunPCA (hnsccScaled, features = VariableFeatures(object=hnsccScaled))
```
Print the data and some plots
```{r print}
print(hnsccScaled[["pca"]],dims=1:5, nfeatures=5)
```
```{r aplot}
VizDimLoadings(hnsccScaled, dims=1:2, reduction="pca")
```
```{r anotherplot}
DimPlot(hnsccScaled, reduction="pca")
```
```{r aheatmap}
DimHeatmap(hnsccScaled, dims=1,balanced=TRUE)
```
```{r anotherheatmap}
DimHeatmap(hnsccScaled, dims=1:15,balanced=TRUE)
```
```{r codenotworking}
#Determine the dimensionality of the dataset
#non si riesce a fare (ucciso)
#hnsccScaled <- JackStraw (hnsccScaled, num.replicate = 100)
#hnsccScaled <- ScoreJackStraw (hnsccScaled, dims = 1:20)
#
#JackStrawPlot(hnsccScaled, dims = 1:15)
```
```{r ElbowPlot}
ElbowPlot(hnsccScaled)
```
##Clustering
We then start with the clustering
```{r clustering}
#clustering
hnsccScaled <- FindNeighbors(hnsccScaled, dims = 1:10)
hnsccScaled <- FindClusters(hnsccScaled, resolution = 0.5)
head(Idents(hnsccScaled),5)

#run umap
hnsccScaled <- RunUMAP (hnsccScaled, dims=1:10)
DimPlot(hnsccScaled, reduction="umap")
```
Save the RDS
```{r save first one}
saveRDS(hnsccScaled, file= '/home/ressf/Documenti/RessBachelorsThesisCode/Downstream_analysis/HNSCC/hnsccScaled.rds')
```

Find the markers of the clusters
```{r cluster 1}
cluster1.markers <- FindMarkers(hnsccScaled, ident.1 = 1, min.pct = 0.25)
head(cluster1.markers, n = 5)
```
```{r cluster 5}
cluster5.markers <- FindMarkers(hnsccScaled, ident.1 = 5, ident.2 = c(0, 3), min.pct = 0.25)
head(cluster5.markers, n = 5)
```
Find markers for every cluster
```{r find markers}
# find markers for every cluster compared to all remaining cells, report only the positive ones
hnsccScaled.markers <- FindAllMarkers(hnsccScaled, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, )
```
```{r groupmarkers}
hnsccScaled.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)
```
Print the table with # cluster and # cells:
```{r table}
table(Idents(hnsccScaled))
```
Create the plot for every cluster:
Cluster 0:
```{r c0}
plotmarker0<- FeaturePlot(hnsccScaled, features=c("S100A2", "S100A9", "KRT16", "ANXA8L1", "KRT14", "S100A8", "GJB2", "COL17A1", "KRT6A", "LGALS7"))
```
Cluster 1:
```{r c1}
plotmarker1<- FeaturePlot(hnsccScaled, features= c("CXCR4", "PTPRC", "CD2", "CD3E", "LCK", "CD3D", "FYN", "SRGN", "GZMK", "CCR7"))
```
Cluster 2:
```{r c2}
plotmarker2<-FeaturePlot(hnsccScaled, features=c("ACTA2", "TAGLN", "MUSTN1", "PLN", "A2M", "SPARCL1", "MYL9", "ADAMTS1", "MYH11", "FHL5"))
```
Cluster 3:
```{r c3}
plotmarker3<-FeaturePlot(hnsccScaled, features=c("EPCAM", "GSTA1", "GSTM1", "CES1", "SPP1", "FGFBP2", "OSGIN1", "ALDH3A1", "GPX2", "UGT1A6"))
```
Cluster 4:
```{r c4}
plotmarker4<-FeaturePlot(hnsccScaled, features=c("COL3A1", "COL6A3", "RARRES2", "MXRA8", "COL1A2", "LUM", "DCN", "CTSK", "COL1A1", "MMP2"))
```
Cluster 5:
```{r c5}
plotmarker5<-FeaturePlot(hnsccScaled,features=c("CXCR6", "CD3D", "SIRPG", "CD2", "PRF1", "TIGIT", "ICOS", "LCK", "GNLY", "GZMB"))
```
Cluster 6:
```{r c6}
plotmarker6<-FeaturePlot(hnsccScaled,features=c("CCNB1", "GJB6", "TK1", "PLEK2", "CDC20", "GJB2", "COL17A1", "FGFBP1", "S100A9", "S100A2"))
```
Cluster 7:
```{r c7}
plotmarker7<-FeaturePlot(hnsccScaled,features=c("ELK2AP", "MZB1", "DERL3", "CD79A", "IGLL5", "IGJ", "NCF1", "BLNK", "SLAMF7", "FKBP11"))
```
Cluster 8:
```{r c8}
plotmarker8<-FeaturePlot(hnsccScaled,features=c("PLVAP", "RAMP3", "RNASE1", "CLEC14A", "DARC", "AQP1", "VWF", "CDH5", "CLDN5", "HYAL2"))
```
Cluster 9:
```{r c9}
plotmarker9<-FeaturePlot(hnsccScaled,features=c("CSN3", "FDCSP", "UBD", "TRIT1", "MGST1", "MYCL1", "ODAM", "SULT1E1", "CLDN1", "KRT19"))
```
Cluster 10:
```{r c10}
plotmarker10<-FeaturePlot(hnsccScaled,features=c("APOD", "SFRP2", "DPT", "PODN", "ABCA8", "FBLN2", "MFAP4", "C3", "CXCL12", "CFD"))
```
Cluster 11:
```{r c11}
plotmarker11<-FeaturePlot(hnsccScaled,features=c("GPX2", "AKR1C2", "AKR1C1", "ALDH3A1", "ADH7", "IPMK", "AKR1C3", "SAMD12", "SUZ12P1", "CYP4F11"))
```
Cluster 12:
```{r c12}
plotmarker12<-FeaturePlot(hnsccScaled,features=c("AZGP1", "ZG16B", "CRISP3", "TCN1", "PIP", "STATH", "MUC7", "PRR4", "SLPI", "PIGR" ))
```
Cluster 13:
```{r c13}
plotmarker13<-FeaturePlot(hnsccScaled,features=c("TPSAB1", "TPSB2", "CPA3", "TPSD1", "HDC", "CTSG", "KIT", "HPGDS", "IL1RL1", "MS4A2"))
```
Cluster 14:
```{r c14}
plotmarker14<-FeaturePlot(hnsccScaled,features=c("C1QB", "AIF1", "C1QC", "MS4A6A", "C1QA", "FCER1G", "FCGR3A", "TYROBP", "IGSF6", "FCGR2A" ))
```


[INSERT HERE THE EXPLANATION ABOUT THE MARKERS AND THE CLUSTERS]
After having searched for the markers we rename the clusters with the name of the cells. We create an object with the clusters' names:
```{r clustersname}
new.cluster.ids <-c("Basal cells", "T helper cells","Ectoderm","Basal cells, maybe cancer basal cells", "Fibroblast", "T cells", "Basal cells in phase S/M", "B cells", "Endothelial cells", "Mesoderm cells", "Fibroblast", "Aorta Endothelial cell", "Airway epithelial cell", "mast cell + neutrophils", "gamma delta T cells")
```
We rename the clusters
```{r rename}
names(new.cluster.ids) <- levels(hnsccScaled)
hnsccScaled <- RenameIdents(hnsccScaled, new.cluster.ids)
```
We plot the clusters with the names
with no legend:
```{r finalplotnol}
DimPlot(hnsccScaled, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```
and with legend:
```{rfinalplotleg}
DimPlot(hnsccScaled, reduction = "umap", label = TRUE, pt.size = 0.5)
```
We save the final object
```{r savefinal}
saveRDS(hnsccScaled, file = "/home/ressf/Documenti/RessBachelorsThesisCode/Downstream_analysis/HNSCC/hnscc_final.rds")
```
